Full-core analysis of thorium-fueled Molten Salt Breeder Reactor using the SERPENT 2 Monte Carlo code

### 2017-08-14 full-core_msbr_model.tex

Zone II is Zone II 37\% salt by volume and each element has a fuel channel diameter of 
6.604cm. It is divided into two different zones: Zone II-A and Zone II-B. The 
graphite elements for Zone II-A are prismatic. Zone II-B elements are 
rectangular slats spaced far enough apart to provide the 0.37 fuel salt 
volume fraction. Fig.~\ref{fig:zone2B} additionally shows the 5.08cm-wide 
annular space between the core graphite and the radial reflector graphite. The 
annulus contains 100\% fuel salt and serves to reduce the damage flux at the 
internal surface of the graphite reflector blocks. The reactor Zone II-B 
graphite 5.08cm-thick slats vary in the radial dimension (average width is 
26.67cm) but are reconstructed without any approximation. From 
the ORNL report \cite{robertson_conceptual_1971}, the suggested design of Zone 
II-B has 8 irregularly-shaped graphite elements every 45$^\circ$ as well as 
salt channels. These graphite elements were simplified into right-circular 
cylindrical shapes with central channels. This is the only simplification made 
to the \gls{MSBR} conceptual geometry in this work.
\FloatBarrier
\begin{figure}[htp!] % replace 't' with 'b' to force it to be on the bottom
  \centering
  \includegraphics[width=\linewidth]{figure_2_4.png}
  \caption{Zone I (left) and Zone II-A (right) elements.}
  \label{fig:zone12A}
\end{figure}
\subsection{Material composition}
The fuel salt, the reactor graphite, and the modified HastelloyN are materials 
unique of the \gls{MSBR} and were created at \gls{ORNL}. The initial fuel salt 
loading composition is LiF-BeF$_2$-ThF$_4$-$^{233}$UF$_4$ (71.8-16-12-0.2 mole 
\%). The lithium in the molten salt fuel is a fully enriched $^{7}$Li because 
$^{6}$Li is a very strong neutron poison and becomes tritium upon neutron 
capture. For cross section generation, ENDF/B-VII was employed \cite{chadwick_endf/b-vii.0:_2006}. The specific temperature was fixed for each material to correctly model the Doppler-broadening of resonance peaks when Serpent generate problem-oriented nuclear data library.
%The isotope composition of each material at the initial state was described in detail in the MSBR conceptual design study \cite{robertson_conceptual_1971} and has been applied to Serpent model without any modification. %
\begin{figure}[htbp!] % replace 't' with 'b' to force it to be on the bottom
  \centering
  \includegraphics[width=0.96\linewidth]{figure_2_5.png}
  \caption{Plan view that includes Zone I, II-A, and II-B elements.}
  \label{fig:zone2B}
\end{figure}
\FloatBarrier
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{R}
This section presents calculation results, such as the effective multiplication 
factor for whole core, neutron flux spectrum, and temperature reactivity 
coefficients. The normalized neutron flux distribution is calculated for the 
whole core using continuous-energy nuclear data.
The temperature coefficients for both fuel salt solution and reactor graphite are computed by comparing effective multiplication factors for two temperatures in the working range.

@@

The fuel temperature coefficient (FTC) is negative due to thermal Doppler 
broadening of the resonance capture cross sections in the thorium and is in a 
good agreement with early research 
\cite{robertson_conceptual_1971,park_whole_2015}. The moderator temperature 
coefficient is positive due to changing density and would increase during reactor operation because of spectrum hardening along with fuel depletion \cite{park_whole_2015}. Finally, the total temperature coefficient of reactivity is relatively large and negative, despite graphite components, and affords excellent reactor stability and controllability.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}
We performed \gls{MSBR} full-core analysis using the Serpent 2 Monte Carlo code. The complex geometry of the reactor is reconstructed in three-dimensional space without any major approximations. Accurate material data was employed to calculate reactor key design parameters. The effective multiplication factor for initial fuel composition is slightly higher than 1 (1.00389) which allows reactor operation from startup to first online reprocessing cycle. The neutron flux energy spectrum was calculated for the whole core and represents the epithermal spectrum of the MSBR. The total temperature coefficient is negative, consequently, the MSBR has negative temperature feedback, but MTC is negative which has a negligible effect on safety because it is outweighed by the strong, negative FTC.

>>>>

Zone II is Zone II 37\% salt by volume and each element has a fuel channel 
diameter of 6.604cm. It is divided into two different zones: Zone II-A and Zone 
II-B. The graphite elements for Zone II-A are prismatic. Zone II-B elements are 
rectangular slats spaced far enough apart to provide the 0.37 fuel salt volume 
fraction. Fig.~\ref{fig:zone2B} additionally shows the 5.08cm-wide annular 
space between the core graphite and the radial reflector graphite. The annulus 
contains 100\% fuel salt and serves to reduce the damage flux at the internal 
surface of the graphite reflector blocks. The reactor Zone II-B graphite 
5.08cm-thick slats vary in the radial dimension (average width is 26.67cm) but 
are reconstructed without any approximation. From the ORNL report 
\cite{robertson_conceptual_1971}, the suggested design of Zone II-B has 8 
irregularly-shaped graphite elements every 45$^\circ$ as well as salt channels. 
These graphite elements were simplified into right-circular cylindrical shapes 
with central channels. This is the only simplification made to the \gls{MSBR} 
conceptual geometry in this work.
\FloatBarrier
\begin{figure}[htp!] % replace 't' with 'b' to force it to be on the bottom
  \centering
  \includegraphics[width=\linewidth]{figure_2_4.png}
  \caption{Zone I (left) and Zone II-A (right) elements.}
  \label{fig:zone12A}
\end{figure}
\subsection{Material composition}
The fuel salt, the reactor graphite, and the modified HastelloyN are materials 
unique of the \gls{MSBR} and were created at \gls{ORNL}. The initial fuel salt 
loading composition is LiF-BeF$_2$-ThF$_4$-$^{233}$UF$_4$ (71.8-16-12-0.2 mole 
\%). The lithium in the molten salt fuel is a fully enriched $^{7}$Li because 
$^{6}$Li is a very strong neutron poison and becomes tritium upon neutron 
capture. For cross section generation, ENDF/B-VII was employed 
\cite{chadwick_endf/b-vii.0:_2006}. The specific temperature was fixed for each 
material to correctly model the Doppler-broadening of resonance peaks when 
Serpent generate problem-oriented nuclear data library.
%The isotope composition of each material at the initial state was described in 
%detail in the MSBR conceptual design study \cite{robertson_conceptual_1971} 
%and has been applied to Serpent model without any modification. %
\begin{figure}[htbp!] % replace 't' with 'b' to force it to be on the bottom
  \centering
  \includegraphics[width=0.96\linewidth]{figure_2_5.png}
  \caption{Plan view that includes Zone I, II-A, and II-B elements.}
  \label{fig:zone2B}
\end{figure}
\FloatBarrier
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{R}
This section presents calculation results, such as the effective multiplication 
factor for whole core, neutron flux spectrum, and temperature reactivity 
coefficients. The normalized neutron flux distribution is calculated for the 
whole core using continuous-energy nuclear data.
The temperature coefficients for both fuel salt solution and reactor graphite 
are computed by comparing effective multiplication factors for two temperatures 
in the working range.

@@

The fuel temperature coefficient (FTC) is negative due to thermal Doppler 
broadening of the resonance capture cross sections in the thorium and is in a 
good agreement with early research 
\cite{robertson_conceptual_1971,park_whole_2015}. The moderator temperature 
coefficient is positive due to changing density and would increase during 
reactor operation because of spectrum hardening along with fuel depletion 
\cite{park_whole_2015}. Finally, the total temperature coefficient of 
reactivity is relatively large and negative, despite graphite components, and 
affords excellent reactor stability and controllability.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}
We performed \gls{MSBR} full-core analysis using the Serpent 2 Monte Carlo 
code. The complex geometry of the reactor is reconstructed in three-dimensional 
space without any major approximations. Accurate material data was employed to 
calculate reactor key design parameters. The effective multiplication factor 
for initial fuel composition is slightly higher than 1 (1.00389) which allows 
reactor operation from startup to the first online reprocessing cycle. The 
neutron flux energy spectrum was calculated for the whole core and represents 
the epithermal spectrum of the MSBR. The total temperature coefficient is 
negative, consequently, the MSBR has negative temperature feedback, but MTC is 
negative which has a negligible effect on safety because it is outweighed by 
the strong, negative FTC.
